### 2022-03-13 paper/localpsf.tex
major edits to beginning of intro. intro beginning section is acceptable now

Our method assumes that the supports of the $\impulseresponse_x$ are contained (or approximately contained) in localized regions. Our method also assumes that the integral kernel is non-negative, i.e., $\Phi(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$. Because of the non-negativity assumption, we are able to estimate the supports of the functions $\phi_x$ a-priori by applying $\Aop$ to a small number of polynomial functions. This, in combination with the locality assumption, allows us to compute impulse responses in batches. We recover the impulse responses $\phi_{x_i}$ for a large number of points $x_i$ with a small number of operator actions $v \mapsto \Aop v$. We also leverage the non-negativity assumption to perform a more sophisticated and accurate interpolation of the impulse responses.

The output of our method is a hierarchical matrix (H-matrix) approximation, $\mathbf{A}_H$, of a discretization of $\Aop$. One may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$ that cannot easily be done using $\Aop$ directly, because $\Aop$ is only available matrix-free. These operations include matrix-matrix addition, matrix-matrix multiplication, (faster) matrix-vector products, matrix factorization, and matrix inversion. H-matrix methods are fast and scale well to large problems. The ability to perform these operations permits, for example, fast solution of Newton linear systems in optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.

>>>>

Our method assumes that the supports of the $\impulseresponse_x$ are contained (or approximately contained) in localized regions. Our method also assumes that the integral kernel is non-negative, i.e., $\Phi(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$. Because of the non-negativity assumption, we are able to estimate the supports of the functions $\phi_x$ a-priori by applying $\Aop$ to a small number of polynomial functions. This, in combination with the locality assumption, allows us to compute impulse responses in batches. We recover the impulse responses $\phi_{x_i}$ for a large number of points $x_i$ with a small number of operator actions $u \mapsto \Aop u$. We also leverage the non-negativity assumption to perform a more sophisticated and accurate interpolation of the impulse responses.

The output of our method is a hierarchical matrix (H-matrix) approximation, $\mathbf{A}_H$, of a discretization of $\Aop$. One may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$ that cannot easily be done using $\Aop$, because $\Aop$ is only available matrix-free. These operations include matrix-matrix addition, matrix-matrix multiplication, (faster) matrix-vector products, matrix factorization, and matrix inversion. H-matrix methods are fast and scale well to large problems. The ability to perform these operations permits, for example, fast solution of Newton linear systems in optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
